---
title: "SQL"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Filtering Columns and Rows
In SQL columns and rows can both be filtered. Columns are filtered with the SELECT command and rows are filtered with the WHERE command. Let's try using the Lahman database.
First, load the Lahman and SQLDF packages.
```{r, message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```
Next, enter the desired filters into a query.
```{r}
query<-"SELECT playerID, teamID, HR, yearID FROM Batting WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Here are some additional examples with different parameters.
Example 1: Shows players, team, and year with at least 40 home runs
```{r}
query<-"SELECT playerID, HR, teamID, yearID FROM Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```
Example 2: Displays players, team, and year with more than 40 home runs but less than 60 strike outs
```{r}
query<-"SELECT playerID, HR, SO, teamID, yearID FROM Batting WHERE HR>40 and SO<60"
sqldf(query)
```
Example 3: Shows Phillies players and year with more than 30 home runs in the 1970's
```{r}
query<-"SELECT playerID, teamID, HR, yearID FROM Batting WHERE teamID='PHI' and HR>30 and yearID<1980 and yearID>=1970"
sqldf(query)
```


#Group By
Example 1
```{r}
query<- "SELECT playerID, SO, AB, teamID FROM Batting WHERE SO<10 and AB>400 ORDER BY SO"
sqldf(query)
```

#Aggrigation
Example 1: Combine column by sum
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting WHERE playerID='ruthba01' GROUP BY playerID"
sqldf(query)
```
Example 2: Filtering with Groups: use HAVING after GROUP instead of WHERE in equations where there is no grouping
```{r}
query<-"SELECT playerID, sum(HR), teamID FROM Batting GROUP BY playerID HAVING sum(HR)>=600 ORDER BY sum(HR) DESC"
sqldf(query)
```
Some other ways to group data are as follows: avg() max() min()

Example 3
```{r}
query<-"SELECT playerID, avg(HR) FROM Batting GROUP BY playerID HAVING avg(HR)>30 ORDER BY avg(HR) DESC"
sqldf(query)
```

#Joins
Example 1: 
Seperated by command for clarity.
```{r}
query <-"SELECT nameFirst, nameLast, name, Batting.yearID, Batting.HR
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```


Example 2
```{r}
query<-"SELECT nameFirst, nameLast, sum(HR)
FROM Batting INNER JOIN Master ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>600"
sqldf(query)
```


Example 3
```{r}
query<-"SELECT nameFirst, nameLast, avg(salary)
FROM Salaries INNER JOIN Master ON Salaries.playerID=Master.playerID
GROUP BY Master.playerID
HAVING avg(salary)>1000000 and nameFirst='Bob'"
sqldf(query)
```

Example 4
```{r}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.teamID 
FROM (Batting inner join Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearId=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Example 5
```{r}
query<-"SELECT nameFirst, nameLast, sum(HR) FROM Batting
INNER JOIN Master on Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) desc"

sqldf(query)
```

Example 6
```{r}
query<-"SELECT nameFirst, nameLast, avg(salary) FROM Salaries
INNER JOIN Master ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) desc"

sqldf(query)
```

